
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 9: Visualization - Revealing a graph"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment9.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

**Notice:** *In this assignment, all the plot should have title, caption, and axis labels. *

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 
```{r}
library(ggplot2)
library(gganimate)
library(gifski)
library(tidyverse)
library(knitr)
library(lubridate)

df <- read_csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')
df$year = year(df$Date_reported)

d1 <- df %>% 
  filter(year == '2024') %>% 
  group_by(Country) %>% 
  summarise(sum = sum(Cumulative_deaths)) %>% 
  mutate(rank=rank(-sum)) 
  
d2 <- d1 %>% filter(rank <= 3) %>% arrange(rank)
```

Hint: 

-  Use `filter` to select the most current date 

-  Pipe with `arrange(-Cumulative_deaths)` to rank the data by the number of cumulative deaths


2. Use `transition_reveal` to make a plot revealing the number of deaths in the three countries with the most numbers of deaths
```{r,cache=TRUE}
df %>% 
  filter(Country %in% c('United States of America', 'Brazil', 'India')) %>% 
  ggplot(aes(y=Cumulative_deaths,
            x=Date_reported,
            color=Country))+ 
  geom_line()+
  geom_point(size=3)+
  geom_text(aes(label = Cumulative_deaths), 
            hjust = -.1, size=5) +
  transition_reveal(Date_reported)
```

3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 
```{r}
df$death_per_cases <- (df$Cumulative_deaths/df$Cumulative_cases)

d1 <- df %>% filter(!Cumulative_cases == '0') %>%  group_by(Country) %>% summarise(Total_death_per_case_Avg = sum(death_per_cases))
d2 <- d1 %>% mutate(rank=rank(-Total_death_per_case_Avg)) 
d3 <- d2 %>% filter(rank <= 3)

d3 %>% 
  arrange(-Total_death_per_case_Avg)

# 1)France, 2)Mauritania, 3)Yemen
```

4. Use `transition_reveal` to make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 
```{r,cache=TRUE}
df %>% 
  filter(!Cumulative_cases == '0', Country %in% c('United States of America', 'Italy', 'Mexico')) %>%
  group_by(Country) %>% 
  ggplot(aes(y=death_per_cases,
            x=Date_reported,
            color=Country))+ 
  geom_line()+
  geom_point(size=3)+
  geom_text(aes(label = death_per_cases), 
            hjust = -.1, size=5) +
  transition_reveal(Date_reported)

```

5. Import the following data https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv
```{r}
df2 <- read.csv('https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv')
df2 <- as_tibble(df2)
df2$date <- as.Date(df2$date)
```
                      
Use `transition_reveal` to plot the total vaccinations per hundred by level of income. Hint: the levels of income are identified in the `location` variable. Put a short comment on the caption of the plot. 
```{r, cache=TRUE, warning=FALSE}
#There were way to many countries to try to do so I filtered out for some I thought would have different results. 

df2 %>% 
  filter(location %in% c('Haiti', 'Kenya','Mexico', 'Norway', 'United States')) %>% 
  group_by(location) %>% 
  ggplot(aes(y=total_vaccinations_per_hundred,
            x=date,
            color= location))+ 
  geom_line()+
  geom_text(aes(label = total_vaccinations_per_hundred), 
            hjust = -.1, size=5) +
  transition_reveal(date)
```

