
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 3.  Working with Data Frame. Base R Style"
---

***Note*:** *This assignment practices working with Data Frame using Base R.*

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment3.Rmd)) in Rstudio. 

- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 3. 

-------

Problems
-------
 \

1. Create the following data frame
```{r}
df = data.frame(Rank = c(0,1,2,3),
                Age = c(28,34,29,42),
                Name = c('Tom', 'Jack', 'Steve', 'Ricky')
                )

df
```


| Rank 	| Age 	| Name  	|
|:------	|:-----	|:-------	|
| 0    	| 28  	| Tom   	|
| 1    	| 34  	| Jack  	|
| 2    	| 29  	| Steve 	|
| 3    	| 42  	| Ricky 	|

2. Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv).

2. Show the names of the variables in the data

```{r}
df1 = read.csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')
names(df1)
```


3. How many columns and rows the data have? 
```{r}
str(df1)
```


4. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?
```{r}
sum(is.na(df1))
colSums(is.na(df1))
#There are no states that currently have missing values.
```
5. What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column. 
```{r}
class(df1$date)
#Currently it is a "character" 
str(df1)

#To change this we have to use as.Date()
df1$date= as.Date(df1$date)
str(df1)
```

6. Capitalize the names of all the variables
```{r}
names(df1) <- c("Date", "State", "Fips", "Cases", "Deaths")
str(df1)
```

7. Find the average number of cases per day. Find the maximum cases a day.  
```{r}
 mean(df1$Cases, na.rm=TRUE)
```

8. How many states are there in the data?  
```{r}
length(table(df1$State))
#There are 56 states in this data set.
```

9. Create a new variable `weekdays` to store the weekday for each rows. 
```{r}
library(lubridate)
df1$weekdays = weekdays(df1$Date)
head(df1)
```

10. Create the categorical variable `death2` variable taking the values as follows

- `has_death` if there is a death that day
-  `no_death` if there is no death that day

Find the frequency and relative frequency of `no_death` and `has_death`. 
```{r}
library(dplyr)
df1$death2 = case_when(df1$Deaths == 1 ~ "has_death", 
                          TRUE ~ "no_death")
prop.table(table(df1$death2))
```


11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint:  Use the `summary` function)

```{r}
summary(df1$Deaths)
#First q = 1171
#Second q = 5035
#Third q = 14912
```


12. Create the categorical variable `death3` variable taking the values as follows

- `low_death` if the number of deaths smaller than the 25 percentile (Q1)

- `mid_death` if the number of deaths from Q1 to Q3

- `high_death` if the number of deaths greater than Q3

```{r}
df1$death3 = case_when(df1$Deaths < 1171 ~ "low_death", 
                       df1$Deaths > 1171 & df1$Deaths < 14912 ~ "mid_death", 
                       df1$Deaths > 14912 ~ "high_death")
```


13. Find the average cases in Rhode Island in 2021
```{r}
df2 <- filter(df1, State == 'Rhode Island', year(Date) == 2021)

mean(df2$Cases)
#Mean # of cases in Rhode Island is 154438.5

```

14. Find the median cases by weekdays in Rhode Island in 2021 
```{r}
by(df2$Cases, df2$weekdays, mean)
```

15. Compare the median cases in Rhode Island in June, July, August and September in 2021. 
```{r}
# Create a variable month
library(lubridate)
df2$month <- month(df2$Date, label = TRUE)


# group months into fewer categories
df2$month2 <- case_when(df2$month %in% c('Jun','Jul','Aug','Sep')~ 'Jun-Sept')

```

16.  Find your own dataset, import it and implement the following functions on the data
```{r}
setwd("C:/Users/student/OneDrive - Bryant University/Desktop/Stats w R/MATH421/Data")
df <-  read.csv('titanic_train.csv')
head(df)
str(df)
names(df)
mean(df$Fare)
min(df$Fare)
max(df$Fare)
table(df$Fare)
sum(is.na(df))
colSums(is.na(df))
class(df)
cor(df$Survived, df$Age)
by(df$Fare, df$Cabin, mean)
df$Fare_Class <- ifelse(df$Fare < 45, 
                        'Cheap_Ticket','Expensive_Ticket') 
df$Age_category = case_when(df$Age < 20 ~ "young", 
                       df$Age > 21 & df$Age < 45 ~ "middle_aged", 
                       df$Age > 46 ~ "old")

```

- head
- str
- names
- mean, min, max
- table
- is.na
- colSums
- class
- cor
- by
- ifelse
- case_when

17.  In the dataset in #16, practice the follows.  You can reuse the code of 16. 

- Create a categorical variable from a continuous variable
- From one categorical variable, create a new categorical variable with fewer categories

If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](../data/titanic.csv)

```{r}
str(df)
#Same from Above
df$Age_category = case_when(df$Age < 20 ~ "young", 
                       df$Age > 21 & df$Age < 45 ~ "middle_aged", 
                       df$Age > 46 ~ "old")
df$Lively = case_when(df$Age_category == "young" &
                        df$Age_category =="middle_aged" ~ "lively",
                      df$Age_category == "old" ~ "not_lively")

```

