---
title: "5_missing_value_notes"
format: html
editor: visual
---

## 5_Missing_value_notes

-   When do we need to handle missing values

    -   When you know they are missing

    -   when building predictive models

    -   Other cases?

-   Checking the NAs

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
df <- read_csv('https://bryantstats.github.io/math421/data/titanic_missing.csv')
```

```{r}
sum(is.na(df))
```

```{r}
#Missing value columns
colSums(is.na(df))
```

```{r, eval=FALSE}
is.na(df)
```

-   **Steps to Handle missing values**

    -   Step 1: Identify the forms of Missing Values (Unknown, Missing, 9999)

    -   Step 2: Convert them to NA

    -   Step 3: Handle it

        -   Approach 1: Remove rows that have any NAs

        -   Approach 2: Fill NAs with the previous or next value of the column

        -   Approach 3: Replace by the mean, majority or a predicted value

```{r}
# Checking Missing Data that uses another word than just "N/A"
colSums(df=="Unknown", na.rm = TRUE)
```

```{r}
# Way to convert back to "N/A"
df2 <- replace(df,df == 'Unkown' |
                  df == 'Missing' |
                  df == 'Not Available', NA)
colSums(df2=="Missing", na.rm = TRUE)

#Another way to do convert to N/A . . . Supposedly not the same as the previous
df3 = replace(df, df %in% c("Unknown", "Missing", "Not Available"), NA)
colSums(df3=="Missing", na.rm = TRUE)
```

```{r}
colSums(df2=="Missing", na.rm = TRUE)

```
