
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 10: Predictive Modeling - Part 1"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment10.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Github Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Use the `Adult Census Income` dataset.  We will predict the income (whether or not it is more than 50k or not) of an adult. Import the dataset.  Partition the data into 80% training and 20% testing.  
```{r}
library(ggplot2)
library(gganimate)
library(gifski)
library(tidyverse)
library(knitr)
library(lubridate)
library(caret)
library(rpart)
library(rattle)
library(bitops)
library(randomForest)

df <- read_csv('https://bryantstats.github.io/math421/data/adult_census.csv')

# Set the target variable
df <- df %>% rename(target=income)

# Correct variables' types
df <- df %>% 
  mutate(target = as.factor(target),
         workclass = as.factor(workclass),
         education = as.factor(education),
         marital.status = as.factor(marital.status),
         occupation = as.factor(occupation),
         relationship = as.factor(relationship),
         race = as.factor(race),
         sex = as.factor(sex),
         native.country = as.factor(native.country)
         )

#Drop all NA rows
sum(is.na(df)) 
# Equals 0 so we don't have to drop any

#Split the data to training and testing
set.seed(2020)
splitIndex <- createDataPartition(df$target, p = .80, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```

2. Practice Decision Tree.  Do the follows:

  - Use `rpart` package, create a decision tree with maximum depth of 3. 
  
  - Calculate the accuracy of the model on the testing data. Notice that the positive outcome here is not `1` but `>50K` or `<50K`. 
  
  - Plot the tree
  
  - Plot the variable importance by the tree
```{r}
#Create a tree
tree_model <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 3))

fancyRpartPlot(tree_model)

#Variable Importances
tree_model$variable.importance
barplot(tree_model$variable.importance)

#Evaluate the Tree
#predict on testing data
pred <- predict(tree_model, df_test, type = "class")
#Evaluate the predictions
cm <- confusionMatrix(data = pred, reference = df_test$target)#Didn't like positive = "1"
cm$overall[1]
#Accuracy = 0.8387592
```

3. Create 3 more trees and compare the testing accuracy of these trees, which tree give the highest testing accuracy.
```{r}
#Tree 1, 4 Leaf
#Create a tree
tree_model1 <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 4))

fancyRpartPlot(tree_model1)

#Variable Importances
tree_model1$variable.importance
barplot(tree_model1$variable.importance)

#Evaluate the Tree 
#predict on testing data
pred1 <- predict(tree_model1, df_test, type = "class")
#Evaluate the predictions
cm1 <- confusionMatrix(data = pred1, reference = df_test$target)#Didn't like positive = "1"
cm1$overall[1]
#Accuracy = 0.8387592

#Tree 2, 5 Leaf
#Create a tree
tree_model2 <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 5))

fancyRpartPlot(tree_model2)

#Variable Importances
tree_model2$variable.importance
barplot(tree_model2$variable.importance)

#Evaluate the Tree
#predict on testing data
pred <- predict(tree_model2, df_test, type = "class")
#Evaluate the predictions
cm <- confusionMatrix(data = pred, reference = df_test$target)#Didn't like positive = "1"
cm$overall[1]
#Accuracy = 0.8387592

#Tree 3, 10 Leaf
#Create a tree
tree_model3 <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 10))

fancyRpartPlot(tree_model3)

#Variable Importances
tree_model3$variable.importance
barplot(tree_model3$variable.importance)

#Evaluate the Tree
#predict on testing data
pred <- predict(tree_model3, df_test, type = "class")
#Evaluate the predictions
cm <- confusionMatrix(data = pred, reference = df_test$target)#Didn't like positive = "1"
cm$overall[1]
#Accuracy = 0.8387592
```



4. Practice Random Forest.  Do the follows: 

  - Use `randomForest` package, create a random forest of 1000 trees. 
  
  - Calculate the accuracy of the model on the testing data. 
  
  - Plot the variable importance by the forest
```{r}
#Random Forest
#Requires installing ("randomForest")
forest_model = randomForest(target ~ ., data=df_train, ntree = 1000)
pred <- predict(forest_model, df_test, type = "class")
cm <- confusionMatrix(data = pred, reference = df_test$target) 
cm$overall[1]
#Accuracy = 0.8587224  
importance(forest_model)
```

5. Create 3 more forests and compare the testing accuracy of these forests, which forest give the highest testing accuracy.
```{r}
#Forest 1
forest_model1 = randomForest(target ~ ., data=df_train, ntree = 1250)
pred <- predict(forest_model1, df_test, type = "class")
cm <- confusionMatrix(data = pred, reference = df_test$target) 
cm$overall[1]
#Accuracy = 0.8599509 
importance(forest_model1)


#Forest 2
forest_model2 = randomForest(target ~ ., data=df_train, ntree = 500)
pred <- predict(forest_model2, df_test, type = "class")
cm <- confusionMatrix(data = pred, reference = df_test$target) 
cm$overall[1]
#Accuracy = 0.8614865 
importance(forest_model2)


#Forest 3
forest_model3 = randomForest(target ~ ., data=df_train, ntree = 200)
pred <- predict(forest_model3, df_test, type = "class")
cm <- confusionMatrix(data = pred, reference = df_test$target) 
cm$overall[1]
#Accuracy =  0.8605651 
importance(forest_model3)
```

6. What is the best model (in term of testing accuracy) among all models (including trees and forests) you have trained?

- Out of all the models including both trees and forests combined the best model proved to be Forest 2 with 500 trees because it had the highest testing accuracy of 86.14865%. 

